Appendix A: Stochastic simulations of noisy phenomena

### 2017-11-01 appendixA/appendixA.Rmd
submitted for review

```{r include=FALSE}
## note: needs library Cairo
knitr::opts_chunk$set(message=FALSE, cache = TRUE, warning = FALSE,  dev="CairoPDF")
```

```{r libraries}
library(nimble)
library(RcppRoll)
library(regimeshifts) ## devtools::install_github("cboettig/regimeshifts")
library(adaptivetau)
library(tidyverse)
library(ggthemes)
library(gridExtra)
colours <- ptol_pal()(2)
```

```{r echo=FALSE}
## plot themes, non-essential. Comment out if theme creates trouble with fonts on your system.
library(Cairo)
library(hrbrthemes)
library(extrafont)
#extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
hrbrthemes::import_roboto_condensed()   # One-time use to set up fonts

@@

```{r gillespie, warning=FALSE, fig.width=7, fig.cap="Population dynamics from a Gillespie simulation of the Levins model with large (N=1000, panel A) and small (N=100, panel B) number of sites (blue) show relatively weaker effects of demographic noise in the bigger system. Models are otherwise identical, with e = 0.2 and c = 1 (code in appendix A). Theoretical predictions for mean and plus/minus one standard deviation shown in horizontal re dashed lines."}
read_csv("../appendixA/gillespie.csv", col_types = "cdiddd")  %>% 
  mutate(system_size = recode(system_size, 
                              large = "A. 1000 total sites", 
                              small= "B. 100 total sites")) %>%
  ggplot(aes(x = time)) + 
  geom_hline(aes(yintercept = mean), lty=2, col=colours[2]) + 
  geom_hline(aes(yintercept = minus_sd), lty=2, col=colours[2]) + 
  geom_hline(aes(yintercept = plus_sd), lty=2, col=colours[2]) + 
  geom_line(aes(y = n), col=colours[1]) +
  facet_wrap(~system_size, scales = "free_y") 
```

@@

```{r quasicycles, fig.width=7, fig.cap="Panel A: under negligible environmental noise, the introduction of a carrying capacity damps Lotka-Volterra oscillations which eventually settle to a stable state.  Panel B: small noise induces quasi-cycles in the same model: sustained long-term oscillations predominately around the natural frequency of the original damped oscillations. Parameters for models shown are r = .1, K = 5,  b = .1, c = .1, d = .1, N = 800; code in Appendix A."}
read_csv("quasicycles.csv") %>%
  gather(species, population, -t, -sigma) %>%
  mutate(sigma = as.factor(sigma)) %>%
  mutate(sigma = recode(sigma, "1e-05" = "A.", "0.01" = "B.")) %>%
  rename(time = t) %>%
  ggplot(aes(time, population, col = species)) + geom_line() + 
  facet_wrap(~ sigma, ncol=2) + scale_color_ptol() 
```

@@

```{r noisy_switch, fig.width=7, fig.height=7, fig.cap="Stochastic switiching in May's model of alternative stable state dynamics.  r = .5, K = 2, Q = 5, H = .38, sigma = .04, a = 0.245, N = 10000"}

@@

```{r tipping, fig.width=7, fig.height=7, fig.cap="Early warning signs of a critical transition in a simulation of increasing serial dilution of budding yeast.  Panel A: Rolling variance   autocorrelation Time series is shown in panel C. Model details and simulation parameters in Appendix A."}
tipping <- read_csv("../appendixA/tipping.csv")

>>>>

```{r include=FALSE}
## note: needs library Cairo
knitr::opts_chunk$set(message=FALSE, cache = 1, warning = FALSE,  dev="CairoPDF")
```

```{r libraries, cache=FALSE}
library(nimble)
library(RcppRoll)
library(regimeshifts) ## devtools::install_github("cboettig/regimeshifts")
library(adaptivetau)
library(tidyverse)
library(ggthemes)
library(gridExtra)
colours <- ptol_pal()(2)
```

```{r echo=FALSE, cache=FALSE}
## plot themes, non-essential. Comment out if theme creates trouble with fonts on your system.
library(Cairo)
library(hrbrthemes)
library(extrafont)
#extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
hrbrthemes::import_roboto_condensed()   # One-time use to set up fonts

@@

```{r gillespie, warning=FALSE, fig.width=7}
read_csv("../appendixA/gillespie.csv", col_types = "cdiddd")  %>% 
  mutate(system_size = recode(system_size, 
                              large = "A. 1000 total sites", 
                              small= "B. 100 total sites")) %>%
  ggplot(aes(x = time)) + 
  geom_hline(aes(yintercept = mean), lty=2, col=colours[2]) + 
  geom_hline(aes(yintercept = minus_sd), lty=2, col=colours[2]) + 
  geom_hline(aes(yintercept = plus_sd), lty=2, col=colours[2]) + 
  geom_line(aes(y = n), col=colours[1]) +
  facet_wrap(~system_size, scales = "free_y") 
```

@@

```{r quasicycles, fig.width=7}
read_csv("quasicycles.csv") %>%
  gather(species, population, -t, -sigma) %>%
  mutate(sigma = as.factor(sigma)) %>%
  mutate(sigma = recode(sigma, "1e-05" = "A.", "0.01" = "B.")) %>%
  rename(time = t) %>%
  ggplot(aes(time, population, col = species)) + geom_line() + 
  facet_wrap(~ sigma, ncol=2) + scale_color_ptol() 
```

@@

```{r noisy_switch, fig.width=7, fig.height=7}

@@

```{r tipping, fig.width=7, fig.height=7}
tipping <- read_csv("../appendixA/tipping.csv")
